Making Sequential Programs Deterministic

### 2019-04-03 java-design.tex
No qualifier polymorphism on classes
small changes

@@

We noticed the following properties with respect to determinism regarding the most widely used \<Map> implementations:
\begin{itemize}
    \item \<HashMap> can never guarantee fully deterministic iteration over its entries. A \codeid{Det HashMap} does not exist.
    \item \<LinkedHashMap> can have any of \<Det>, \<OrderNonDet>, or \<NonDet> types. Iterating over a \<LinkedHashMap> returns
    its entries in the order of their insertion. An \<OrderNonDet LinkedHashMap> can be created by calling its constructor with a \<HashMap>.
    \item \<TreeMap> can either be \<Det> or \<NonDet>. An \<OrderNonDet TreeMap> doesn't exist because the entries always ordered.
\end{itemize}

>>>>

\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
\todo{We do not support qualifier polymorphism on classes, only on methods.}
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forallt{\kappa} \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

@@

We noticed the following properties with respect to determinism regarding the most widely used \<Map> implementations:
\begin{itemize}
    \item \<HashMap> can never guarantee fully deterministic iteration over its entries. A \codeid{Det HashMap} does not exist.
    \item \<LinkedHashMap> can have any of \<Det>, \<OrderNonDet>, or \<NonDet> types. Iterating over a \<LinkedHashMap> returns
    its entries in the order of their insertion. An \<OrderNonDet LinkedHashMap> can be created by calling its constructor with a \<HashMap>.
    \item \<TreeMap> can either be \<Det> or \<NonDet>. An \<OrderNonDet TreeMap> doesn't exist because the entries are always ordered.
\end{itemize}
